Digital History in Finland, 1488-1828

### 2017-01-31 inst/examples/manuscript-krakow.Rmd
analysis updates

  res <- plot_timeline(dfs, field = "titlecount", group = "gatherings", nmin = 15, mode = "absolute") 
  print(res + ggtitle(paste(catal, " (n)", sep = "")))

  res <- plot_timeline(dfs, field = "titlecount", group = "gatherings", nmin = 15, mode = "relative") 
  print(res + ggtitle(paste(catal, " (%)", sep = "")))
}
```

@@

>>>>

  res <- plot_timeline(data = dfs, x = "publication_year", y = "titlecount", group = "gatherings", nmin = 15, mode = "absolute") 
  print(res + ggtitle(paste(catal, " (n)", sep = "")))

  res <- plot_timeline(data = dfs, x = "publication_year", y = "titlecount", group = "gatherings", nmin = 15, mode = "relative") 
  print(res + ggtitle(paste(catal, " (%)", sep = "")))
}
```

@@

### Paper consumption per title

Paper consumption per title in 1757–1765, 1766–1774, and 1775–1783 in
Stockholm, Lund and Uppsala in Kungliga.

```{r paper_per_title, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=4}
catalogue <- "Kungliga"
places <- c("Stockholm", "Lund", "Uppsala")
periods <- list(c(1757, 1765), c(1766, 1774), c(1775, 1783))

df <- df0
df$period <- rep(NA, nrow(df))
for (i in 1:length(periods)) {
  pr <- periods[[i]]
  df$period[df$publication_year >= min(pr) & df$publication_year <= max(pr)] <- paste(min(pr), max(pr), sep = "-")
}
df$period <- factor(df$period)

# Selected catalogue with selected years
df <- filter(df, catalog == catalogue & publication_place %in% places & !is.na(period) & !is.na(paper))

df <- df %>% group_by(period, publication_place) %>%
             summarise(n = n(), paper = sum(paper, na.rm = TRUE)) %>%
	     mutate(paper_per_title = paper/n)

theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_place, y = paper_per_title)) +
     geom_bar(stat = "identity", position = "dodge", aes(fill = period), color = "black") +
     ylab("Paper per title (sheets)") +
     xlab("Publication period") +
     #scale_color_manual(values = c("blue", "darkgreen")) + 
     # 	 scale_size_discrete(range = c(2, 8)) +
     guides(fill = guide_legend(reverse = TRUE, title = "Place"), color = "none") 
print(p)

kable(df)
```

### Book production and Riksdag assemblies

According to Kungliga in the the long eighteenth century.

```{r riksdar, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}
minyear <- 1690
maxyear <- 1828 # or 1810
catal <- "Kungliga"

library(magrittr)
library(ggplot2)
library(dplyr)
library(ggplot2)
library(reshape2)

  df <- df0
  df <- subset(df, catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year <= maxyear)

  li <- list(1719, # Stockholm 20 januari 1719 1 juni 1719
             1720,
	     1723,
	     c(1726, 1727),
	     1731,
	     1734, # Stockholm 14 maj 1734 14 december 1734
	     c(1738, 1739),
	     c(1740, 1741),
	     1742,
	     c(1746, 1747),
	     c(1751, 1752),
	     c(1755, 1756),
	     c(1760, 1762),
	     c(1765, 1766), # Stockholm 21 februari 1765 21 oktober 1766
	     c(1769, 1770), # Norrköping & Stockholm 22 april 1769 5 februari 1770
	     c(1771, 1772), # Stockholm 19 juni 1771 12 september 1772
	     c(1778, 1779),
	     1786,
	     1789,
	     1792, # Gävle 26 januari 1792 24 februari 1792         
	     1800,
	     c(1809, 1810))

  rect_left <- c(min(na.omit(dfm$date)))
  for (item in li) {
    rect_left <- c(rect_left, min(item) - .5, max(item) + .5)
  }
  rect_left <- c(rect_left, max(na.omit(dfm$date)))

  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("Background", "Highlight"), length = nrow(rectangles))

 cols <- c("lightgray", "darkgray")
 bgcol <- "white"
 rectangles$shade[rectangles$shade == "Background"] <- bgcol
 rectangles$shade[rectangles$shade == "Highlight"] <- rep(cols, length.out = sum(rectangles$shade == "Highlight"))
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
       aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=shade), alpha=0.8) + 
         scale_fill_manual(values = levels(rectangles$shade)) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  #p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 10))
  p <- p + scale_x_continuous(breaks=scales::pretty_breaks(n = 20))  
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(catal)
  p <- p + guides(fill = "none") 
  print(p)
```

### Book production

Book production by year in Turku in Fennica and Kungliga 1640–1828 as a percentage of all books in the catalogues.

```{r bookprod, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
minyear <- 1640
maxyear <- 1828 
df <- df0 %>% filter(!is.na(publication_place)) 
df$myplace <- df$publication_place
df$myplace[!df$myplace == "Turku"] <- "Other"

df <- df %>%  filter(publication_decade >= minyear & publication_decade <= maxyear) %>%
	      group_by(catalog, publication_decade, myplace) %>%
	      summarize(n = n())

pics <- list()
for (cat in unique(df$catalog)) {
  dfs <- subset(df, catalog == cat)
  p <- ggplot(dfs, aes(x = publication_decade, y = n)) +
       geom_bar(aes(fill = myplace), color = "black", stat = "identity", position = "stack") +
       scale_fill_manual(values = c("gray", "black")) +
       ylab("Title count") +
       xlab("Publication decade") +
       guides(fill = guide_legend(title = "Place")) +
       ggtitle(cat) +
       xlim(values = c(min(df$publication_decade), max(df$publication_decade)))
  pics[[cat]] <- p
}

library(gridExtra)
grid.arrange(pics[[1]], pics[[2]], nrow = 2)
```

### Kungliga / Linkoping

```{r linkoping, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
minyear <- 1640
maxyear <- 1828
myplace <- "Linköping"
cat <- "Kungliga"
df <- df0 %>% filter(publication_place == myplace) %>%
      	      filter(catalog == cat) %>%
              filter(publication_year >= minyear &
	             publication_year <= maxyear)

tab <- unique(df[, c("publication_year", "author", "title", "publisher")])
write.table(tab, sep = "\t", quote = F, row.names = FALSE, file = "output.tables/Linkoping.tab")
```

[Document listing for `r paste(cat, "/", myplace, minyear, "-", maxyear, sep = " ")`](output.tables/Linkoping.tab)

### Title count in Linköping

```{r linkoping2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
p <- plot_timeline(data = df, x = "publication_decade", y = "titlecount") +
       ylab("Title count (n)") +
       ggtitle(paste("Title count: ", cat, "/", myplace))
print(p)
.```

### Unique publishers in Linköping by year

Never two publishers seen on the same year.

```{r linkoping3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
dfs <- df %>% select(publication_year, publisher) %>%
	      filter(!is.na(publisher)) %>%
              unique %>%	      
	      arrange(publication_year) 
kable(dfs)
.```

### Title count per capita

The historical population sizes used in this analysis are shown in this [table](https://github.com/rOpenGov/bibliographica/blob/master/inst/extdata/population_sizes_in_cities.csv).

```{r percapita, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Read historical population sizes
f <- system.file("inst/extdata/population_sizes_in_cities.csv", package = "bibliographica")
pop <- read.csv(f)
pop <- melt(pop, "Year")
colnames(pop) <- c("publication_decade", "publication_place", "population")
pop <- as_data_frame(pop)
pop$publication_place <- as.character(pop$publication_place)

# Cities with population data
cities <- setdiff(pop$publication_place, "Year")

minyear <- 1640
maxyear <- 1828
cat <- "Kungliga"
df <- df0 %>% filter(publication_place %in% cities) %>%
      	      filter(catalog == cat) %>%
              filter(publication_decade %in% pop$publication_decade) %>%
	      select(publication_decade, publication_place) %>%	      
	      group_by(publication_decade, publication_place) %>%
	      summarise(n = n())

# Remove special chars for compatibility (table join fails otherwise)
pop$publication_place <- gsub("Linköping", "Linkoping", pop$publication_place)
df$publication_place <- gsub("Linköping", "Linkoping", df$publication_place)

# Join title count and population size
dfm <- inner_join(pop, df, by = c("publication_decade", "publication_place"))

# Add title count per capita
dfm <- dfm %>% mutate(titles_per_capita = n/population)

p <- ggplot(dfm, aes(x = publication_decade, y = titles_per_capita)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication decade") +
       ylab("Title count per capita (n)") +
       ggtitle("Title count per capita in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)       
```
